---
title: "lab3_assembly_annotation"
output: html_document
---

library("IRanges")
library("GenomicRanges")
library("Rsamtools")
library("GenomicAlignments")
library("Gviz")


Assembly statistics:

        Total length:   12679670
        Fragments:      354
        Fragments N50:  52932
        Largest frg:    302366
        Scaffolds:      0
        Mean coverage:  8

#4.1 (Q1)
Although the total length is close to the estimated size of the genome, the results are very fragmented, at 354 fragments/contigs. Also the coverage of the sample is not adequate at 8, as it hints at a low consensus accuracy 

#5.1.1 

```{r4, eval=FALSE}
C:33.9%[S:33.7%,D:0.1%],F:0.4%,M:65.8%,n:3282,E:5.1% 
	1111	Complete BUSCOs (C)	(of which 57 contain internal stop codons)  
	1107	Complete and single-copy BUSCOs (S)	 
	4	Complete and duplicated BUSCOs (D)	   
	13	Fragmented BUSCOs (F)			  
	2158	Missing BUSCOs (M)			  
	3282	Total BUSCO groups searched	
```

Given the very low coverage of this, I would say the results are tragic. BUSCO shows 1/3 level of completeness, which is in line with what I might expect given low coverage.

The BUSCO N50 was close to the N50 from flye, showing 52932 from flye and ~52 kbp from BUSCO

The total assembly length from BUSCO outputs was 12679670, which is exactly the same as the length from flye

#5.2.1

The quast genome fraction is 0.529 (52.9%). Compared to the BUSCO metric which showed a completeness of ~33%, the results don't agree with each other. 

The duplication ratio from quast is 1.156

The level of misassemblies and length of misassemblies both being zero tells me that the reference genome is a great fit for my isolated genome. It also shows that nothing crazy happened to the structure, like translocations or inversions.

#6.2
install.packages("BiocManager")            
BiocManager::install("Biostrings, force=TRUE")    
library(Biostrings)

in_fa  <- "./GCF_000146045.2_R64_genomic.fna"
out_fa <- "R64_chr1.fna"

# Read all sequences
fa <- readDNAStringSet(in_fa)

# Find the entry for chromosome I (handles common header variants)
pick <- grep("(\\bchrI\\b|chromosome I|NC_001133\\.9)", names(fa), ignore.case = TRUE)
pick <- pick[1]


writeXStringSet(fa[pick], out_fa, width = 60)
cat(sprintf("Wrote %s (%d bp)\n", out_fa, width(fa[pick])))


BiocManager::install("Rsamtools",force=TRUE)
library(Rsamtools)
BiocManager::install("GenomicAlignments")
library(GenomicAlignments)
BiocManager::install("Gviz",force=TRUE)
library(Gviz)
# BAM file path
bam_file <- "contig_vs_chr1.sorted.bam"

# Open connection to BAM file
bam <- BamFile(bam_file)

# Read all alignments (already restricted to chr1)
alns <- readGAlignments(bam)

# Convert to GRanges object
gr_alns <- granges(alns)
options(ucscChromosomeNames = FALSE)
aln_track <- AnnotationTrack(gr_alns, name = "Contigs", genome = "sacCer3", chromosome = "chr1")
axis_track <- GenomeAxisTrack()

plotTracks(list(axis_track, aln_track),
           from = min(start(gr_alns)),
           to = max(end(gr_alns)),
           main = "Contig Alignments to Chromosome 1")
           
length(gr_alns)

#6.3.2.1
There are 2 contigs aligning to chr1

```{r plotTracks, eval=TRUE,fig.width=8, fig.height=4}
knitr::include_graphics("plotsTrack_lab3.pdf")
```

#6.3.3
out_fa <- "longest_contig.fasta"
lens <- width(fa)
idx  <- which.max(lens)
writeXStringSet(fa[idx], out_fa, width = 60)
cat(sprintf("Wrote longest contig: %s (%d bp) → %s\n", names(fa)[idx], lens[idx], out_fa))

#6.3.4
minimap2 -ax splice -uf --secondary=no ./longest_contig.fasta GCF_000146045.2_R64_genomic.fna > cds_aln.sam

#6.3.5
samtools view -bS cds_aln.sam > cds_aln.bam

samtools sort cds_aln.bam -o cds_aln.sorted.bam

samtools index cds_aln.sorted.bam

#6.3.6
options(ucscChromosomeNames = FALSE)
bam_file <- "cds_aln.sorted.bam"
bam <- BamFile(bam_file)
alns <- readGAlignments(bam)
gr_alns <- granges(alns)
aln_track <- AnnotationTrack(gr_alns, name = "Contigs", genome = "sacCer3", chromosome = "chr1")
axis_track <- GenomeAxisTrack()
plotTracks(list(axis_track, aln_track),
           from = min(start(gr_alns)),
           to = max(end(gr_alns)),
           main = "Contig Alignments to longest contig")
      
```{r plotTracks2, eval=TRUE,fig.width=8, fig.height=4}
knitr::include_graphics("longest_contig_track.pdf")
```
